 Allocation scoring rules as a generalization of quantile scoring rules

### 2023-06-13 notes/allocation_notes.rmd

Let $F,F_i$ be general distributions such as forecaster beliefs or reports about $Y$, and $G$ be the true distribution of $Y$. 
For a general scoring rule $S$, the $S$-entropy of $F$ is defined as 
\begin{align}
e_S(F) &:= S(F, F) = E_F[S(F, Y)].
\end{align}
Applying this to $S_\alpha$, we can define the $g,\alpha$-entropy of $F$ as
\begin{align}
e_{\alpha}(F) &:= S_{\alpha}(F, F) = E_F[s_{\alpha}(q_{F,\alpha}, Y)] .
\end{align}
If a forecaster believes that $Y \sim F$ then they expect to receive a score of $e_{\alpha}(F)$ by reporting their $\alpha$ quantile.
According to \eqref{eqn:scalar_obj_cdf_form} and \eqref{eqn:obj_pexp},
\begin{align}
e_{\alpha}(F)
&= (1-\alpha) \int_{-\infty}^{q_{\alpha,F}} F(u) dg(u) + \alpha \int_{{q_{\alpha,F}}}^{\infty}(1-F(u)) dg(u) \\
&= \alpha \overline{g}\sb{F}-\overline{g}_{F}(x).
\end{align}
Similarly, we define the $g,\alpha$-divergence between two distributions as
\begin{align}
d_{\alpha}(F_1,F_2) &:= S_{\alpha}(F_1,F_2) - S_{\alpha}(F_2,F_2) = S_{\alpha}(F_1,F_2) - e_{\alpha}(F_2). \label{egn:divergence}
\end{align}
Again from \eqref{eqn:scalar_obj_cdf_form} we get
\begin{align}
d_{\alpha}(F_1,F_2)
&= (1-\alpha) \int_{-\infty}^{q_{\alpha,F_1}} F_2(u) dg(u) + \alpha \int_{{q_{\alpha,F_1}}}^{\infty}(1-F_2(u)) dg(u) \\
&-(1-\alpha) \int_{-\infty}^{q_{\alpha,F_2}} F_2(u) dg(u) - \alpha \int_{{q_{\alpha,F_2}}}^{\infty}(1-F_2(u)) dg(u) \\
&= (1-\alpha) \int_{q_{\alpha,F_2}}^{q_{\alpha,F_1}} F_2(u) dg(u) + \alpha \int_{{q_{\alpha,F_1}}}^{q_{\alpha,F_2}}(1-F_2(u)) dg(u).
\end{align}
And from \eqref{eqn:obj_pexp},
\begin{align}
d_{\alpha}(F_1,F_2) &= g(q_{\alpha,F_1})(F_2(q_{\alpha,F_1}) - \alpha) -\overline{g}_{F_2}(q_{\alpha,F_1}) + \alpha \overline{g}\sb{F_2}  - \alpha \overline{g}\sb{F_2} + \overline{g}_{F_2}(q_{\alpha,F_2}) \\
&= g(q_{\alpha,F_1})(F_2(q_{\alpha,F_1}) - \alpha) + \int_{q_{\alpha,F_1}}^{q_{\alpha,F_2}}g(y)dF_2(y)
\end{align}
To emphasize the reality of $G$ we write $e_{\alpha}(G)$ as $e_{\alpha}(Y)$, which with \eqref{egn:divergence}
gives 
\begin{align}
S_{\alpha}(F,G) = d_{\alpha}(F,G) + e_{\alpha}(Y). \label{eqn:marg_score_decomp}
\end{align}
This expresses the true expected score of $F$ 
(i.e., not that calculated by the forecaster) as the sum of the failure of $F$ to match $G$ and the inherent uncertainty of $Y$.

@@

In the inventory management literature, minimization of an expectation of the form of $\sbar_F(x)$ for a given $F$ arises in the \emph{newsvendor problem}.  This involves the ordering decision faced by a retailer of a perishable good (such as newpapers) when customer demand is uncertain. Because the focus here is on revenues and expenditures, terms other than just the over- and underprediction penalties in \eqref{eqn:scalar_score} often appear, giving a retailer's scoring function as
\begin{align}
s_{\mathbf{n}}(x,y) & = c x - p \min(x,y) - u(x-y)\sb{+} + r(y-x)\sb{+} \label{eqn:news_score}
\end{align}
where $\mathbf{n} = \{c,p,u,r\}$.
This score is netting the monetary transfers that occur when
where $x$ newspapers are ordered at price $c$, $\min(y,x)$ are sold at retail price $p$, $(x-y)\sb{+}$ are sold at salvage price $u$, 
and $r(y-x)\sb{+}$ of monetarily valued customer "goodwill" is lost. 
Using $x = y + (x-y)_+ - (y-x)_+$ and $\min(x,y) = y - (y-x)_+$ \eqref{eqn:news_score} can be rewritten as
\begin{align}
s_{\mathbf{n}}(x,y) &= (c-u)(x-y)\sb{+} + (p-c+r)(y-x)\sb{+} -(p-c)y
\end{align}
giving the retailer's objective function under a demand distribution $F$ as
\begin{align}
\sbar_{F, {\mathbf{n}}}(x) = (c-u)E_F[(x-Y)\sb{+}] + (p-c+r)E_F[(Y-x)\sb{+}] -(p-c)E[Y].
\end{align}
This differs by a constant from our $\sbar_{F,O,U}(x)$ in \eqref{eqn:scalar_obj} with $g(x)=x$, $O=c-u$ 
and $U=p-c+r$, and so has the same minimizer   
\begin{align}
Q = F^{-1}\left(\frac{U}{O+U}\right) = F^{-1}\left(\frac{p-c+r}{p-u + r}\right).
\end{align} 
Note, for reference, that in the literature $\sbar_{F, {\mathbf{n}}}(x)$ is often encountered in the form
\begin{align}
&(O+U) \int_{0}^{x} (x - y)dF(y) - Ux+ (U-u+r)E[Y] \\{}
=\,&(p+r-u) \int_{0}^{x} (x - y)dF(y) - (p-c+r)x+ rE[Y]
\end{align}

>>>>

Let $F,F_i$ be general distributions such as forecaster beliefs or reports about $Y$, and $G$ be the true distribution of $Y$. 
For a general scoring rule $S$, we define the $S$-entropy of $F$ as 
\begin{align}
e_S(F) &:= S(F, F) = E_F[S(F, Y)]
\end{align}
and the $S$-entropy of the random variable $Y$ as
\begin{align}
e_S(Y) &:= e_S(G).
\end{align}
Applying this to $S_\alpha$, we can define the $g,\alpha$-entropy of $F$ as
\begin{align}
e_{\alpha}(F) &:= S_{\alpha}(F, F) = E_F[s_{\alpha}(q_{F,\alpha}, Y)] .
\end{align}
A forecaster believing that $Y \sim F$ will expect to receive a score of $e_{\alpha}(F)$ by reporting their $\alpha$ quantile, while an
omniscient forecaster \emph{knowing} that $Y\sim G$ will expect a score of $e_S(Y)$ for reporting $q_{\alpha,Y}$.
According to \eqref{eqn:scalar_obj_cdf_form} and \eqref{eqn:obj_pexp},
\begin{align}
e_{\alpha}(F)
&= (1-\alpha) \int_{-\infty}^{q_{\alpha,F}} F(u) dg(u) + \alpha \int_{{q_{\alpha,F}}}^{\infty}(1-F(u)) dg(u) \\
&= \alpha \overline{g}\sb{F}-\overline{g}_{F}(q_{\alpha,F}).
\end{align}
Similarly, we define the $g,\alpha$-divergence between two distributions as
\begin{align}
d_{\alpha}(F_1,F_2) &:= S_{\alpha}(F_1,F_2) - S_{\alpha}(F_2,F_2) = S_{\alpha}(F_1,F_2) - e_{\alpha}(F_2). \label{egn:divergence}
\end{align}
Again from \eqref{eqn:scalar_obj_cdf_form} we get
\begin{align}
d_{\alpha}(F_1,F_2)
&= (1-\alpha) \int_{-\infty}^{q_{\alpha,F_1}} F_2(u) dg(u) + \alpha \int_{{q_{\alpha,F_1}}}^{\infty}(1-F_2(u)) dg(u) \\
&-(1-\alpha) \int_{-\infty}^{q_{\alpha,F_2}} F_2(u) dg(u) - \alpha \int_{{q_{\alpha,F_2}}}^{\infty}(1-F_2(u)) dg(u) \\
&= (1-\alpha) \int_{q_{\alpha,F_2}}^{q_{\alpha,F_1}} F_2(u) dg(u) + \alpha \int_{{q_{\alpha,F_1}}}^{q_{\alpha,F_2}}(1-F_2(u)) dg(u).
\end{align}
And from \eqref{eqn:obj_pexp},
\begin{align}
d_{\alpha}(F_1,F_2) &= g(q_{\alpha,F_1})(F_2(q_{\alpha,F_1}) - \alpha) -\overline{g}_{F_2}(q_{\alpha,F_1}) + \alpha \overline{g}\sb{F_2}  - \alpha \overline{g}\sb{F_2} + \overline{g}_{F_2}(q_{\alpha,F_2}) \\
&= g(q_{\alpha,F_1})(F_2(q_{\alpha,F_1}) - \alpha) + \int_{q_{\alpha,F_1}}^{q_{\alpha,F_2}}g(y)dF_2(y)
\end{align}
With \eqref{egn:divergence}, we get 
\begin{align}
S_{\alpha}(F,G) = d_{\alpha}(F,G) + e_{\alpha}(Y), \label{eqn:marg_score_decomp}
\end{align}
expressing the true expected score of $F$ 
(i.e., not that calculated by the forecaster) as the sum of the failure of $F$ to match $G$ and the inherent uncertainty of $Y$.

@@

In the inventory management literature, minimization of an expectation of the form of $\sbar_F(x)$ for a given $F$ arises in the \emph{newsvendor problem}.  This involves the ordering decision faced by a retailer of a perishable good (such as newpapers) when customer demand is uncertain. Because the focus here is on revenues and expenditures, terms other than just the over- and underprediction penalties in \eqref{eqn:scalar_score} often appear, giving a retailer's scoring function as
\begin{align}
s_{\mathbf{n}}(x,y) & = c x - p \min(x,y) - u(x-y)\sb{+} + r(y-x)\sb{+} \label{eqn:news_score} \\
&= \begin{cases}
cx - px + r(y-x), & x \leq y \\
cx - py - u(x-y), & x > y
\end{cases}
\end{align}
where $\mathbf{n} = \{c,p,u,r\}$.
This score is netting the monetary transfers that occur when
where $x$ newspapers are ordered at price $c$, $\min(y,x)$ are sold at retail price $p$, $(x-y)\sb{+}$ are sold at salvage price $u$, 
and the "goodwill", monetarily valued at $r$, of $(y-x)\sb{+}$ customers is lost. 
Using $x = y + (x-y)_+ - (y-x)_+$ and $\min(x,y) = y - (y-x)_+$ \eqref{eqn:news_score} can be rewritten as
\begin{align}
s_{\mathbf{n}}(x,y) &= (c-u)(x-y)\sb{+} + (p-c+r)(y-x)\sb{+} -(p-c)y
\end{align}
giving the retailer's objective function under a demand distribution $F$ as
\begin{align}
\sbar_{F, {\mathbf{n}}}(x) = (c-u)E_F[(x-Y)\sb{+}] + (p-c+r)E_F[(Y-x)\sb{+}] -(p-c)E_F[Y].
\end{align}
This differs by a constant from our $\sbar_{F,O,U}(x)$ in \eqref{eqn:scalar_obj} with $g(x)=x$, $O=c-u$ 
and $U=p-c+r$, and so has the same minimizer   
\begin{align}
q_{F,O,U} = F^{-1}\left(\frac{U}{O+U}\right) = F^{-1}\left(\frac{p-c+r}{p-u + r}\right).
\end{align} 
Note, for reference, that in the literature $\sbar_{F, {\mathbf{n}}}(x)$ is often encountered in the form
\begin{align}
&(O+U) E_F[(x-Y)\sb{+}] + U E_F[Y] E_F[(Y-x)\sb{+}-(x-Y)\sb{+}] -(p-c)E_F[Y] \\
&(O+U) E_F[(x-Y)\sb{+}] + U (E_F[Y] - x) -(p-c)E_F[Y]\\
=\,&(p+r-u) \int_{0}^{x} (x - y)dF(y) - (p-c+r)x+ rE_F[Y]
\end{align}
